---
title: "RNA basis RRP40 KD experiment"
output: html_document
---

```{r load libraries,message=FALSE}
library(scales) # v 1.0.0
```

```{r read in counts}
counts_RNA <- read.table('../Mapping/output/counts_cDNA_RRP40KDscreens.txt',stringsAsFactors=F)
counts_RNA_scaled <- as.data.frame(scale(counts_RNA[-nrow(counts_RNA),],
                                         colSums(counts_RNA), center=F))

counts_gDNA <- read.table('../Mapping/output/counts_gDNA_RRP40KDscreens.txt',stringsAsFactors=F)
counts_gDNA_scaled <- as.data.frame(scale(counts_gDNA[-nrow(counts_gDNA),],
                                          colSums(counts_gDNA), center=F))
```

Put unspliced and spliced versions in side-by-side columns, calculate the totals
```{r Unspliced-spliced side-by-side}
RNA_perinsert <- as.data.frame(matrix(0,nrow=nrow(counts_gDNA_scaled),ncol=ncol(counts_RNA_scaled)*6))
colnames(RNA_perinsert) <-
          paste(rep(colnames(counts_RNA_scaled),each=6),
                c('unspliced',paste0('spliced',1:3),'splicedtotal','total'),sep="_")
rownames(RNA_perinsert) <- rownames(counts_gDNA_scaled)
RNA_perinsert[,paste(colnames(counts_RNA_scaled),'unspliced',sep="_")] <-
  counts_RNA_scaled[1:nrow(counts_gDNA_scaled),]

for(sample in colnames(counts_RNA_scaled)) {
	for(i in 1:3){
		RNA_perinsert[,paste0(sample,'_spliced',i)] <-
			sapply(rownames(RNA_perinsert),function(x){
      				counts_RNA_scaled[paste0(x,'_spliced',i),sample]
    	})
	}
  RNA_perinsert[,paste0(sample,'_splicedtotal')] <- 
    rowSums(RNA_perinsert[,grep(paste0(sample,'_spliced.$'),colnames(RNA_perinsert))],na.rm=T)
  RNA_perinsert[,paste0(sample,'_total')] <-
    rowSums(RNA_perinsert[,c(paste0(sample,'_unspliced'),paste0(sample,'_splicedtotal'))],na.rm=T)
}
```

Normalize to gDNA
```{r Cutoff and norm to gDNA}
# These are thresholds I settled on to balance improving correlation between experiments and not losing too much data.
thresholds <- c("gDNA_siNT_A"=1e-5,"gDNA_siNT_B"=1e-5,"gDNA_siNT_C"=1e-5,
                "gDNA_siRRP40_A"=1e-5,"gDNA_siRRP40_B"=1e-5,"gDNA_siRRP40_C"=1e-5,"gDNA_siINTS11_A"=1e-5) 

RNA_over_gDNA <- RNA_perinsert
for(experiment in c('siNT_A','siNT_B','siNT_C','siRRP40_A','siRRP40_B','siRRP40_C','siINTS11_A')) {
	RNA_over_gDNA[,grep(paste0(experiment,"_"),colnames(RNA_over_gDNA))] <-
	  RNA_over_gDNA[,grep(paste0(experiment,"_"),colnames(RNA_over_gDNA))] /
                       	       counts_gDNA_scaled[,grep(paste0(experiment,"$"),colnames(counts_gDNA_scaled))]
	RNA_over_gDNA[counts_gDNA_scaled[,grep(paste0(experiment,"$"),colnames(counts_gDNA_scaled))] <
	                thresholds[grep(paste0(experiment,"$"),names(thresholds))],
	              grep(paste0(experiment,"_"),colnames(RNA_over_gDNA))] <- NA
}
```

Set median of randomized controls to 1
```{r Norm to randomized controls}
load('input/HV20200429_full_library.RData')
controls <- full_library$unique_id[full_library$category=="randomized_control"]

RNA_over_ctrl <- RNA_over_gDNA
for(sample in colnames(counts_RNA_scaled)){
  RNA_over_ctrl[,grep(sample,colnames(RNA_over_ctrl))] <- 
    RNA_over_gDNA[,grep(sample,colnames(RNA_over_gDNA))] / 
      median(RNA_over_gDNA[controls, paste(sample,'total',sep="_")],na.rm=T)
}
```

Calculate splicing efficiencies
```{r Splicing efficiencies}
splicingefficiencies <- data.frame("siNT_A_splicingefficiency"=
                                     RNA_over_ctrl$cDNA_siNT_A_splicedtotal / RNA_over_ctrl$cDNA_siNT_A_total,
                                   "siNT_B_splicingefficiency"=
                                     RNA_over_ctrl$cDNA_siNT_B_splicedtotal / RNA_over_ctrl$cDNA_siNT_B_total,
                                   "siNT_C_splicingefficiency"=
                                     RNA_over_ctrl$cDNA_siNT_C_splicedtotal / RNA_over_ctrl$cDNA_siNT_C_total,
                                   "siRRP40_A_splicingefficiency"=
                                     RNA_over_ctrl$cDNA_siRRP40_A_splicedtotal / RNA_over_ctrl$cDNA_siRRP40_A_total,
                                   "siRRP40_B_splicingefficiency"=
                                     RNA_over_ctrl$cDNA_siRRP40_B_splicedtotal / RNA_over_ctrl$cDNA_siRRP40_B_total,
                                   "siRRP40_C_splicingefficiency"=
                                     RNA_over_ctrl$cDNA_siRRP40_C_splicedtotal / RNA_over_ctrl$cDNA_siRRP40_C_total
                                   )
RNA_over_ctrl <- cbind(RNA_over_ctrl,splicingefficiencies)
```

Correlations between replicates after all normalizations
```{r Correlations}
cor(RNA_over_ctrl$cDNA_siNT_A_total,RNA_over_ctrl$cDNA_siNT_B_total,use="complete.obs",method="spearman")
cor(RNA_over_ctrl$cDNA_siNT_A_total,RNA_over_ctrl$cDNA_siNT_C_total,use="complete.obs",method="spearman")
cor(RNA_over_ctrl$cDNA_siRRP40_A_total,RNA_over_ctrl$cDNA_siRRP40_B_total,use="complete.obs",method="spearman")
cor(RNA_over_ctrl$cDNA_siRRP40_A_total,RNA_over_ctrl$cDNA_siRRP40_C_total,use="complete.obs",method="spearman")

cor(RNA_over_ctrl$cDNA_siNT_A_total,RNA_over_ctrl$cDNA_siRRP40_A_total,use="complete.obs",method="spearman")
```

Add pseudocounts to everything with < 5e-3 normalized counts
```{r Pseudocounts}
RNA_over_ctrl_pseudo <- RNA_over_ctrl
RNA_over_ctrl_pseudo[,grep('_total',colnames(RNA_over_ctrl_pseudo))][
  RNA_over_ctrl_pseudo[,grep('_total',colnames(RNA_over_ctrl_pseudo))]< 5e-3] <- 5e-3
```

Plots after all normalizations and adding pseudocounts
```{r Correlation plots w pseudocounts}
plot(log(RNA_over_ctrl_pseudo$cDNA_siNT_A_total,2),
     log(RNA_over_ctrl_pseudo$cDNA_siNT_B_total,2),
     pch=16,cex=0.6,col=alpha(1,.3),#xlim=c(-6.5,6),ylim=c(-6.5,6),
     main="Total-RNA siNT A vs B",xlab="siNT - A (log2)",ylab="siNT - B (log2)")
plot(log(RNA_over_ctrl_pseudo$cDNA_siRRP40_A_total,2),
     log(RNA_over_ctrl_pseudo$cDNA_siRRP40_B_total,2),
     pch=16,cex=0.6,col=alpha(1,.3),#xlim=c(-6.5,6),ylim=c(-6.5,6),
     main="Total-RNA siRRP40 A vs B",xlab="siRRP40 - A (log2)",ylab="siRRP40 - B (log2)")
```

Average replicates. Divide by median of controls again after averaging, so that the controls are really set to 1.
Calculate correlations.
Then add pseudocounts and make plots
```{r Average replicates}
RNA_over_ctrl_totals <- RNA_over_ctrl[,grep('_total|splicing',colnames(RNA_over_ctrl))]
RNA_over_ctrl_averages <- RNA_over_ctrl_totals
RNA_over_ctrl_averages$siNT_average <-
    rowMeans(RNA_over_ctrl_averages[,grep('siNT_._total',colnames(RNA_over_ctrl_averages))],na.rm=T)
RNA_over_ctrl_averages$siNT_average <- RNA_over_ctrl_averages$siNT_average /
                                      median(RNA_over_ctrl_averages[controls,"siNT_average"],na.rm=T)
RNA_over_ctrl_averages$siNT_splicingeff_average <-
    rowMeans(RNA_over_ctrl_averages[,grep('siNT_._splicing',colnames(RNA_over_ctrl_averages))],na.rm=T)
RNA_over_ctrl_averages$siRRP40_average <-
    rowMeans(RNA_over_ctrl_averages[,grep('siRRP40_._total',colnames(RNA_over_ctrl_averages))],na.rm=T)
RNA_over_ctrl_averages$siRRP40_average <- RNA_over_ctrl_averages$siRRP40_average /
                                      median(RNA_over_ctrl_averages[controls,"siRRP40_average"],na.rm=T)
RNA_over_ctrl_averages$siRRP40_splicingeff_average <-
    rowMeans(RNA_over_ctrl_averages[,grep('siRRP40_._splicing',colnames(RNA_over_ctrl_averages))],na.rm=T)
RNA_over_ctrl_averages <- RNA_over_ctrl_averages[,grep('average',colnames(RNA_over_ctrl_averages))]

# Correlations
cor(RNA_over_ctrl_averages$siNT_average,RNA_over_ctrl_averages$siRRP40_average,use="complete.obs",method="spearman")

RNA_over_ctrl_averages_pseudo <- RNA_over_ctrl_averages
RNA_over_ctrl_averages_pseudo[,c(1,3)][
                    RNA_over_ctrl_averages[,c(1,3)]< 5e-3] <- 5e-3
plot(log(RNA_over_ctrl_averages_pseudo$siNT_average,10),
     log(RNA_over_ctrl_averages_pseudo$siRRP40_average,10),
     pch=16,cex=0.6,col=alpha(1,.15),#xlim=c(-2.5,2),ylim=c(-2.5,2),
     main="siNT vs siRRP40",xlab="siNT average (log10)",ylab="siRRP40 average (log10)")
```

Save
```{r Save}
save(RNA_over_ctrl,RNA_over_ctrl_pseudo,file="output/normalized_RNA_all_RRP40KDexp.RData")
save(RNA_over_ctrl_averages,RNA_over_ctrl_averages_pseudo,
     file="output/normalized_RNA_averagedtotals_RRP40KDexp.RData")
save(RNA_over_ctrl_totals,file="output/normalized_RNA_totals_RRP40KDexp.RData")
```
